---
title: "learning spatial files"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

install.packages("devtools")
library(tidyverse)
library(urbnmapr)
library(devtools)
library(ggplot2)

devtools::install_github("UrbanInstitute/urbnmapr")

ggplot() + 
  geom_polygon(data = urbnmapr::states, mapping = aes(x = long, y = lat, group = group),
		           fill = "grey", color = "white") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45)

tx_weather$CZ_FIPS = as.character(tx_weather$CZ_FIPS)
class(counties$county_fips)

dim(tx_weather)


counties_tx = counties[counties$state_name == "Texas",]
tx_weather$BEGIN_LAT_R <- round(tx_weather$BEGIN_LAT,1) 
tx_weather$BEGIN_LON_R <- round(tx_weather$BEGIN_LON,1) 

#joining by long/lat just to see (this doesn't really work)
#joined_weather <- left_join(tx_weather, counties_tx, by= c("BEGIN_LON_R" = "long", "BEGIN_LAT_R" = "lat"))

#joined by FIPS code
joined_weather <- left_join(tx_weather, counties_tx, by= c("CZ_FIPS" = "county_fips"))
dim(joined_weather)
joined_weather <- joined_weather[!is.na(joined_weather$BEGIN_LAT),] 

joined_weather %>%
  ggplot(aes(BEGIN_LON, BEGIN_LAT,group = group, if(!all(is.na(DAMAGE_PROPERTY))){fill=DAMAGE_PROPERTY})) +
  geom_polygon(color = NA) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "Property Damage")

# texas_weather code 
joined_weather %>% 
  ggplot(mapping = aes(BEGIN_LON, BEGIN_LAT, group = group, fill = DAMAGE_PROPERTY_NUM)) +
  geom_polygon(color = "#ffffff", size = .25) +
  scale_fill_gradientn(labels = scales::percent,colors=NA,
                       guide = guide_colorbar(title.position = "top")) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")) +
  labs(fill = "Property Damage")


```

#trying again with SF and usaboundaries 
tutorial: https://ryanpeek.org/2019-04-29-spatial-joins-in-R/

```{r}
library(sf)
library(USAboundaries)
devtools::install_github("ropensci/USAboundariesData")
library(USAboundariesData)
library(ggplot2)
library(scales)

us <- us_boundaries(type="state", resolution = "low") %>% 
  filter(!state_abbr %in% c("PR", "AK", "HI"))

tx <- USAboundaries::us_states(resolution = "high", states = "TX")

# make a box around texas (a grid with an n=1) for inset
tx_box <- st_make_grid(tx, n = 1)

# get texas county boundary
tx_co <- USAboundaries::us_counties(resolution = "high", states = "TX")

# make sure we have all the pieces with a quick test plot
plot(tx$geometry, col="gray50", border="maroon")
plot(tx_co$geometry, add=T, border="pink", col=NA)
plot(tx_box, add=T, border="red3", col=NA, lwd=2)

#have to filter out misisng values to make it work 
filt_tx_weather <- tx_weather[!is.na(tx_weather$BEGIN_LAT),] 

# make spatial
filt_tx_weather <- as.data.frame(filt_tx_weather) %>% 
  st_as_sf(coords=c("BEGIN_LON","BEGIN_LAT"), crs=4326, remove=FALSE)  

plot(filt_tx_weather$geometry, cex=0.5)
plot(tx$geometry, col=alpha("gray", 0.5), border="#440154FF", lwd=1.5, add=TRUE)
plot(filt_tx_weather$geometry, add=T, pch=21, bg="#21908CFF", cex=0.7, col="black")
title("NOAA Texas Extreme Weather Events Beginning Locations")

```



# Some Links I've been using 
https://www.maths.lancs.ac.uk/~rowlings/Teaching/UseR2012/cheatsheet.html


```{r}
library(sp)
library(rgdal)
library(rgeos)
library(raster)
library(plyr)
library(dplyr)
library(maps)
library(mapdata)
library(maptools)

cords <- cbind(filt_tx_weather$BEGIN_LON, filt_tx_weather$BEGIN_LAT)
sp_cords_tx <- SpatialPoints(cords) # makes them spatial 
plot(sp_cords_tx) # plot! 
summary(sp_cords_tx) # summarizes cords 
coordinates(sp_cords_tx) # prints cords

crs.geo <-CRS("+init=EPSG:32633")
proj4string(sp_cords_tx) <- crs.geo  # define projection system of our data
is.projected(sp_cords_tx)
summary(sp_cords_tx)
st_crs(filt_tx_weather) <- 4326 # hard coding the CRS compatibility 
filt_tx_weather <- st_transform(filt_tx_weather, 4326)  # hard coding the CRS compatibility 

#merge sp cords back into tx_weather 
sp_tx_df <- SpatialPointsDataFrame(sp_cords_tx, as.data.frame(filt_tx_weather))
summary(sp_tx_df)

#merged_weather <- join(tx_weather, counties_tx, by.x = "CZ_FIPS", by.y ="county_fips")
cords <- cbind(counties_tx$long, counties_tx$lat)
sp_cords <- SpatialPoints(cords) # makes them spatial
crs.geo <-CRS("+init=EPSG:32633")
proj4string(sp_cords) <- crs.geo  # define projection system of our data
is.projected(sp_cords)
summary(sp_cords)
sp_tx_counties <- SpatialPointsDataFrame(sp_cords, as.data.frame(counties_tx))
sf_counties <- st_as_sf(sp_tx_counties)
sf_counties <- st_as_sfc(counties_tx)
sf_counties <- st_transform(sf_counties, 32633)  # hard coding the CRS compatibility 

###this did not work, retrying 
#weather_merge <- over(filt_tx_weather, sf_counties)
#weather_merge <- sapply(st_intersects(sp_tx_df,sf_counties), function(z) if (length(z)==0) NA_integer_ else z[1])
#weather_recombined <- cbind(sp_tx_df, weather_merge)
#plot(weather_recombined)
#plot(weather_recombined$DAMAGE_PROPERTY_NUM)
```
```{r}
# Why does this not work? Overlay
plot(tx_counties_sf) # ¥ay made a map of counties of texas 
par(new = T)
plot(sp_tx_df, col = "blue", add = T) # events
par(new = F)

par(mfrow = c(1, 2))
plot(tx_counties_sf, axes = TRUE)
plot(sp_tx_df, axes = TRUE)
#clearly not on the same projection
```

OVER Merge 
```{r}
# making spatial polygons files 
counties_sf <- get_urbn_map("counties", sf = TRUE)
counties_shape_file <- as_Spatial(counties_sf, IDs = counties_sf$county_fips)
#counties_shape_file <- as(counties_shape_file, "Spatial") # makes a SpatialPolygons object 
# Spatial Polygons DataFrame
tx_counties_sf <- subset(counties_shape_file, counties_shape_file$state_abbv == "TX")

# setting CRS 
proj4string(sp_tx_df) <- crs.geo
proj4string(tx_counties_sf) <- crs.geo

#tx_counties_sf <- CRS(proj4string(tx_counties_sf))
#tx_events_proj <- spTransform(sp_tx_df, tx_counties_sf)

```


trying to get the same CRS 
```{r}
# this is a spatial polygons dataframe 
head(tx_counties_sf)

# county map needs to be re-projected to be long/lat 
tx_counties_sf <- sp_transform(tx_counties_sf, sp_tx_df)
crs.geo <- CRS("+init=EPSG:4326")
proj4string(tx_counties_sf) <- crs.geo
is.projected(tx_counties_sf)
summary(tx_counties_sf)
```

```{r}
library(maptools)
#counting num of events in each county
###over.list <- over(tx_counties_sf, geometry(sp_tx_df), returnList = TRUE)
over.list <- over(geometry(sp_tx_df), tx_counties_sf, returnList = TRUE)
merged.weather <- spCbind(sp_tx_df, over.list)
num.events <- sapply(over.list, length)
counties <- spCbind(tx_counties_sf, num.events)

events_per_dist <- over(counties, merged.weather, returnList = FALSE)
events_per_dist[!is.na(events_per_dist)]

```


# Maybe if I used sf package instead? 
```{r}
# sf type counties data 
class(counties_sf)
plot(counties_sf)

ct_tx_sf <- counties_sf[counties_sf$state_abbv == 'TX',]

plot(ct_tx_sf)

sf_tx_df = st_as_sf(sp_tx_df) # this warning should be okay

#set CRS the same 
sf_tx_df <- st_set_crs(sf_tx_df, 2163)

# this join doesn't seem to work
joined <- st_join(sf_tx_df, ct_tx_sf)

# this doesn't work either 
intersects <- st_intersects(sf_tx_df, ct_tx_sf)

# doesn't work, can't merge over two sf objects
joined <- merge(sf_tx_df, ct_tx_sf, by.x= "CZ_FIPS", by.y="county_fips")

# checking if there's any overlap
any(st_touches(sf_tx_df, ct_tx_sf, sparse = FALSE))
# doesn't seem to be any 'touches'

# trying a non-overlaping join and Re-Projecting
sf_tx_df_P = st_transform(sf_tx_df, 4326)
ct_tx_sf_P = st_transform(ct_tx_sf, 4326)
extent(ct_tx_sf_P)

sel = st_is_within_distance(sf_tx_df_P, ct_tx_sf_P, dist = 20)
summary(lengths(sel) > 0) # this did not seem to work at all whatsoever 

```

# Stepping back 

I am really trying to take the weather data and the lat/long associated with it to then find the county_fips code. This should be a spatial join on the long/lat and then grabbing the county_fips and adding that to the tx_weather file. 

# trying again 
```{r}
# x=points, y=polygon
pts <- sf_tx_df_P$geometry
pol <- ct_tx_sf_P$geometry
joined <- st_contains(pts, pol)
mat = st_intersects(pts, pol, sparse = FALSE)
```

# Found the issue 
Even with CRS set to the same and etc, I am realizing that the coordinate values have the wrong PERIOD! Instead of being after the first two digits, the coordinate is after the first 6 digits! Why!!! Now I just need to Gsub and I should be back on track. 


```{r}
ct_tx_sf_mod <- ct_tx_sf
bad_cords <- as.data.frame(st_coordinates(ct_tx_sf))
bad_cords[1,"X"] 

bad_cords[1]
# these g subs are not right but they're what I have 
bad_cords$X <- as.numeric(gsub("([[:digit:]]{6,6})$", ".\\1", 
                      bad_cords$X)) 
bad_cords$Y <- as.numeric(gsub("([[:digit:]]{6,6})$", ".\\1", 
                      bad_cords$Y)) 

```
This might not be the issue -- ask Paulina tomorrow. 


```{r}
sf_tx_df_P = st_transform(sf_tx_df, 4326)
ct_tx_sf_P = st_transform(ct_tx_sf, 4326)
plot(sf_tx_df_P)
plot(ct_tx_sf_P)

lst <- st_intersects(ct_tx_sf_P, sf_tx_df_P) # empty 
mat <- st_intersects(ct_tx_sf_P, sf_tx_df_P, sparse=FALSE) # empty 
apply(mat, 1, any)
lengths(lst) > 0

st_intersects(ct_tx_sf_P, sf_tx_df_P)[[1]] # nothing n first polygon 

ggplot() +
  #geom_sf(data = sf_tx_df_P$geometry, col="blue") + 
  geom_sf(data = ct_tx_sf_P) +
  ggtitle("Map of Weather Events in Texas Counties") +
  coord_sf()

ggplot() +
  geom_sf(data = sf_tx_df_P$geometry, col="blue") + 
  #geom_sf(data = ct_tx_sf_P) +
  ggtitle("Map of Weather Events in Texas Counties") +
  coord_sf()


sf_cords_tx <- st_as_sf(sp_cords_tx)

joined <- st_intersects(sf_cords_tx, sf_counties)
joined <- st_intersects(sf_cords_tx, sf_counties,sparse=FALSE) # empty 


sel = st_is_within_distance(sf_cords_tx, sf_counties, dist = 20)
summary(lengths(sel) > 0) # THIS WORKED!!!!!! OH my lord 

trial <- sapply(st_intersects(sf_cords_tx,tx_co), function(z) if (length(z)==0) NA_integer_ else z[1])
trial <- merge(sf_cords_tx, sf_counties)


plot(sf_counties)
plot(sf_cords_tx, add = TRUE) 

plot(sf_cords_tx)
plot(sf_counties['county_fips'], add = TRUE) 
# well this is something? maybe? 

plot(sf_counties['county_fips'], main="Weather Events Grouped by County Fips Code", type="p")
plot(sf_cords_tx, add = TRUE, pch=19, type="l") 

#plot(tx_co, type= 'l')
plot(sf_counties, type='p')

###over.list <- over(tx_counties_sf, geometry(sp_tx_df), returnList = TRUE)
over.list <- over(geometry(sp_tx_df), tx_counties_sf, returnList = TRUE)
merged.weather <- spCbind(sp_tx_df, over.list)
num.events <- sapply(over.list, length)
counties <- spCbind(tx_counties_sf, num.events)

events_per_dist <- over(counties, merged.weather, returnList = FALSE)
events_per_dist[!is.na(events_per_dist)]


```


```{r}
sf_cords_tx = st_transform(sf_cords_tx, 4326)


joined <- st_intersects(sf_cords_tx, tx_co)
joined_sparse <- st_intersects(sf_cords_tx, tx_co, sparse=TRUE)
